---
output:
  pdf_document: default
  html_document: default
---


# Data

We will use the same dataset as Project 1: [`movies_merged`](https://s3.amazonaws.com/content.udacity-data.com/courses/gt-cs6242/project/movies_merged).

# Objective

Your goal in this project is to build a linear regression model that can predict the `Gross` revenue earned by a movie based on other variables. You may use R packages to fit and evaluate a regression model (no need to implement regression yourself). Please stick to linear regression, however.

# Instructions

You should be familiar with using an [RMarkdown](http://rmarkdown.rstudio.com) Notebook by now. Remember that you have to open it in RStudio, and you can run code chunks by pressing *Cmd+Shift+Enter*.

Please complete the tasks below and submit this R Markdown file (as **pr2.Rmd**) containing all completed code chunks and written responses, as well as a PDF export of it (as **pr2.pdf**) which should include all of that plus output, plots and written responses for each task.

_Note that **Setup** and **Data Preprocessing** steps do not carry any points, however, they need to be completed as instructed in order to get meaningful results._

# Setup

Same as Project 1, load the dataset into memory:

```{r}
load('movies_merged')
```

This creates an object of the same name (`movies_merged`). For convenience, you can copy it to `df` and start using it:

```{r}
df = movies_merged
cat("Dataset has", dim(df)[1], "rows and", dim(df)[2], "columns", end="\n", file="")
colnames(df)
```

## Load R packages

Load any R packages that you will need to use. You can come back to this chunk, edit it and re-run to load any additional packages later.

```{r}
library(ggplot2)
library(glmnet)
```

If you are using any non-standard packages (ones that have not been discussed in class or explicitly allowed for this project), please mention them below. Include any special instructions if they cannot be installed using the regular `install.packages('<pkg name>')` command.

**Non-standard packages used**: None

# Data Preprocessing

Before we start building models, we should clean up the dataset and perform any preprocessing steps that may be necessary. Some of these steps can be copied in from your Project 1 solution. It may be helpful to print the dimensions of the resulting dataframe at each step.

## 1. Remove non-movie rows

```{r}
# TODO: Remove all rows from df that do not correspond to movies
#let's see if any NA in df Type column and what types in the column.
sum(is.na(df$Type))
table(df$Type)
#There is no NA, and in total 40000 rows of movies. Let's remove all non-movie rows from df.
df=df[df$Type=='movie',]
cat("After removal, dataset has", dim(df)[1], "rows", end="\n", file="")
```

## 2. Drop rows with missing `Gross` value

Since our goal is to model `Gross` revenue against other variables, rows that have missing `Gross` values are not useful to us.

```{r}
# TODO: Remove rows with missing Gross value
#check how many rows with NA for Gross column
sum(is.na(df$Gross))
#remove all rows with NA in Gross
df=df[!is.na(df$Gross),]
#confirm the remaining rows
dim(df)
```

## 3. Exclude movies released prior to 2000

Inflation and other global financial factors may affect the revenue earned by movies during certain periods of time. Taking that into account is out of scope for this project, so let's exclude all movies that were released prior to the year 2000 (you may use `Released`, `Date` or `Year` for this purpose).

```{r}
# TODO: Exclude movies released prior to 2000
#check how many rows with Year 2000 or later
sum(df$Year>=2000)
#exclude movies prior to 2000
df=df[df$Year>=2000,]
#confirm the remaining rows
dim(df)
```

## 4. Eliminate mismatched rows

_Note: You may compare the `Released` column (string representation of release date) with either `Year` or `Date` (numeric representation of the year) to find mismatches. The goal is to avoid removing more than 10% of the rows._

```{r}
# TODO: Remove mismatched rows
#extract year information from Released variable
RLyear=as.numeric(format(df$Released,'%Y'))
#remove rows that df$Year and df$Released have more than 1 year difference
df=df[abs(df$Year-RLyear)<=1,]
#confirm more than 90% rows left after removing mismatches.
dim(df)
```

## 5. Drop `Domestic_Gross` column

`Domestic_Gross` is basically the amount of revenue a movie earned within the US. Understandably, it is very highly correlated with `Gross` and is in fact equal to it for movies that were not released globally. Hence, it should be removed for modeling purposes.

```{r}
# TODO: Exclude the `Domestic_Gross` column
df$Domestic_Gross=NULL
```

## 6. Process `Runtime` column

```{r,warning=F}
# TODO: Replace df$Runtime with a numeric column containing the runtime in minutes
# first let's see how many time formats in the Runtime column
runtime=df$Runtime
sum(is.na(runtime))
rtFormat=as.factor(gsub("[0-9]+ *","",runtime))
summary(rtFormat)
#make Runtime to numeric value in mins
df$Runtime=sapply(runtime,function(str) as.numeric(gsub("min", "", str)))
```

Perform any additional preprocessing steps that you find necessary, such as dealing with missing values or highly correlated columns (feel free to add more code chunks, markdown blocks and plots here as necessary).

```{r}
# TODO(optional): Additional preprocessing
df=na.omit(df)
#remove Date since it is highly related to Year
df$Date=NULL
#remove three tomato review related columns since they are highly related to some of #the other tomato review columns 
df$tomatoUserMeter=NULL
df$tomatoMeter=NULL
df$tomatoReviews=NULL

```

_**Note**: Do NOT convert categorical variables (like `Genre`) into binary columns yet. You will do that later as part of a model improvement task._

## Final preprocessed dataset

Report the dimensions of the preprocessed dataset you will be using for modeling and evaluation, and print all the final column names. (Again, `Domestic_Gross` should not be in this list!)

```{r}
# TODO: Print the dimensions of the final preprocessed dataset and column names
print(paste("df has",dim(df)[1],"rows and",dim(df)[2],"columns"))
colnames(df)
```

# Evaluation Strategy

In each of the tasks described in the next section, you will build a regression model. In order to compare their performance, use the following evaluation procedure every time:

1. Randomly divide the rows into two sets of sizes 5% and 95%.
2. Use the first set for training and the second for testing.
3. Compute the Root Mean Squared Error (RMSE) on the train and test sets.
4. Repeat the above data partition and model training and evaluation 10 times and average the RMSE results so the results stabilize.
5. Repeat the above steps for different proportions of train and test sizes: 10%-90%, 15%-85%, ..., 95%-5% (total 19 splits including the initial 5%-95%).
6. Generate a graph of the averaged train and test RMSE as a function of the train set size (%).

You can define a helper function that applies this procedure to a given model and reuse it.

# Tasks

Each of the following tasks is worth 20 points. Remember to build each model as specified, evaluate it using the strategy outlined above, and plot the training and test errors by training set size (%).

## 1. Numeric variables

Use linear regression to predict `Gross` based on all available _numeric_ variables.

```{r}
# TODO: Build & evaluate model 1 (numeric variables only)
dfn=df[,c("Gross","Budget","tomatoUserReviews","tomatoUserRating","tomatoRotten","tomatoRating","tomatoFresh","imdbVotes","imdbRating", "Runtime","Year")]
N=dim(dfn)[1]
#helper function to iterate  10 times for computing RMSEs given a train size.
get_rmse_list=function(df,prop_to_train,N){
  #lists to record computations
  rmse_list=c()
  prop_to_train_list=c()
  type=c()
  #set random seeds for good comparisons between tasks.
  set.seed(200)
  seeds=sample(10000,10)
  #ten iterations
  for( seed in seeds){
    set.seed(seed)
    #train test partitions
    to_train=sample(N,prop_to_train*N)
    train=df[to_train,]
    test=df[-to_train,]
    #build models using train data
    model=lm(Gross~.,data = train)
    #make prediction  on train data by model
    pred_trainGross=predict(model,train[,-1])
    #compute train error
    rmse_train=sqrt(mean((pred_trainGross-train[,1])^2))
    #make prediction on test data by model
    pred_testGross=predict(model,test[,-1])
    #compute test error
    rmse_test=sqrt(mean((pred_testGross-test[,1])^2))
    #record the results
    rmse_list=c(rmse_list,rmse_train,rmse_test)
    prop_to_train_list=c(prop_to_train_list,prop_to_train,prop_to_train)
    type=c(type,"train","test")
  }
  data.frame(prop_to_train=prop_to_train_list,type=type,rmse=rmse_list)
}

#start computation from train size 5%
rmse_df=get_rmse_list(dfn,0.05,N)
#iteration from train size 10% to 95%
for (prop_to_train in (2:19)*0.05){
  rmse_df=rbind(rmse_df,get_rmse_list(dfn,prop_to_train,N))
}
#summary rmse to get mean and se
average_rmse=aggregate(rmse_df["rmse"],by=rmse_df[c("type","prop_to_train")],function(X) c(mean=mean(X),sd=sd(X),N=length(X)))
average_rmse=do.call(data.frame,average_rmse)
names(average_rmse)[3:5]=c("rmse","sd","N")
average_rmse$se=average_rmse$sd/sqrt(average_rmse$N)

#plot mean and se of RMSEs for each train size
ggplot(average_rmse,aes(prop_to_train,rmse,group=type,col=type))+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin=rmse-se,ymax=rmse+se),width=0.02)+
  ylab("RMSE")+
  theme(axis.text.x = element_text(angle=90, hjust=1))+
  ggtitle("Task 1. RMSE over in and out of sample by train size")
```

**Q**: List all the numeric variables you used.

**A**: 

Here are the numeric variables I used in this question.

"Budget","tomatoUserReviews","tomatoUserRating","tomatoRotten",
"tomatoRating","tomatoFresh","imdbVotes",
"imdbRating", "Runtime","Year".

And use them to make linear regression model for Gross. 

So in summary, after task 1, the average train errors are around 96 M, and the average test errors are around 90 M.

```{r}
#print average train rmse
average_rmse$rmse[average_rmse$type=="train"]
#print average test rmse 
average_rmse$rmse[average_rmse$type=="test"]
```


## 2. Feature transformations

Try to improve the prediction quality from **Task 1** as much as possible by adding feature transformations of the numeric variables. Explore both numeric transformations such as power transforms and non-numeric transformations of the numeric variables like binning (e.g. `is_budget_greater_than_3M`).


```{r,warning=FALSE}
#sqrt transformation five numeric variables including Gross
dfn2=sqrt(dfn[c("Gross","Budget","tomatoRotten","tomatoFresh","imdbVotes")])

#binning Runtime
qruntime=q=c(-Inf,quantile(dfn$Runtime,probs = seq(0.1,.9,.1),na.rm = T),Inf)
runtimelist=paste0("Runtime",1:10)
dfn2$Runtime=cut(dfn$Runtime,qruntime,labels =runtimelist )
for(runtime in runtimelist){
  runtimeCol=data.frame(as.numeric(grepl(runtime,dfn2$Runtime)))
  names(runtimeCol)=runtime
  dfn2=cbind(dfn2,runtimeCol)
}


#binning imdbrate
qimdbrate=q=c(-Inf,quantile(dfn$imdbRating,probs = seq(0.1,.9,.1),na.rm = T),Inf)
imdbratelist=paste0("imdbRating",1:10)
dfn2$imdbRating=cut(dfn$imdbRating,qimdbrate,labels =imdbratelist )
for(imdbrate in imdbratelist){
  imdbrateCol=data.frame(as.numeric(grepl(imdbrate,dfn2$imdbRating)))
  names(imdbrateCol)=imdbrate
  dfn2=cbind(dfn2,imdbrateCol)
}


#add the other four numerical variables

dfn2=cbind(dfn2,df[c("tomatoRating","tomatoUserRating","tomatoUserReviews","Year")])
dfn2o=dfn2
#remove factor columns Runtime and imdbRating, and only remain binary columns
dfn2$Runtime=NULL 
dfn2$imdbRating=NULL

#let's see if need remove any column
summary(lm(Gross~.,dfn2))
#since one out of all Runtime classes can be calculated by another 9
#same thing applies to imdbRating columns
#we have to remove Runtime9 and imdbRating9 columns
dfn2[,c(14,24)]=NULL

N=dim(dfn2)[1]
#helper function to iterate  10 times for computing RMSEs given a train size.
get_rmse_list2=function(df,prop_to_train,N){
  set.seed(200)
  seeds=sample(10000,10)
  rmse_list=c()
  prop_to_train_list=c()
  type=c()
  for( seed in seeds){
    set.seed(seed)
    #train and test partition
    to_train=sample(N,prop_to_train*N)
    train=df[to_train,]
    test=df[-to_train,]
    #build model
    model=lm(Gross~.,data = train)
    #make predictions and compute RMSEs
    pred_trainGross=(predict(model,train[,-1]))^2
    rmse_train=sqrt(mean((pred_trainGross-(train[,1])^2)^2))
    
    pred_testGross=(predict(model,test[,-1]))^2
    rmse_test=sqrt(mean((pred_testGross-(test[,1])^2)^2))
    
    #record results
    rmse_list=c(rmse_list,rmse_train,rmse_test)
    prop_to_train_list=c(prop_to_train_list,prop_to_train,prop_to_train)
    type=c(type,"train","test")
  }
  data.frame(prop_to_train=prop_to_train_list,type=type,rmse=rmse_list)
}
#iterate train size to build models and compute RMSEs 
rmse_df2=get_rmse_list2(dfn2,0.05,N)
for (prop_to_train in (2:19)*0.05){
  rmse_df2=rbind(rmse_df2,get_rmse_list2(dfn2,prop_to_train,N))
}
#summary RMSEs to get mean and se
average_rmse2=aggregate(rmse_df2["rmse"],by=rmse_df2[c("type","prop_to_train")],function(X) c(mean=mean(X),sd=sd(X),N=length(X)))
average_rmse2=do.call(data.frame,average_rmse2)
names(average_rmse2)[3:5]=c("rmse","sd","N")
average_rmse2$se=average_rmse2$sd/sqrt(average_rmse2$N)

#plot the results from above summary
ggplot(average_rmse2,aes(prop_to_train,rmse,group=type,col=type))+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin=rmse-se,ymax=rmse+se),width=0.02)+
  ylab("RMSE")+
  theme(axis.text.x = element_text(angle=90, hjust=1))+
  ggtitle("Task 2. RMSE over in and out of sample by train size")
```


**Q**: Explain which transformations you used and why you chose them.

**A**: 
In this section, I did sqrt transformation on "Gross","Budget","tomatoRotten","tomatoFresh","imdbVotes". Since after sqrt transformation, the relationship between Gross and the others is much closer to linear. See plots below:

```{r}
plot(Gross~Budget,dfn2)
plot(Gross~tomatoRotten,dfn2)
plot(Gross~tomatoFresh,dfn2)
plot(Gross~imdbVotes,dfn2)
```

I also binned imdbRating and Runtime to 10 classes for each by combining quantile and cut functions.
The reasons for binning is that before binning, we didn't see any strong linear relationships, even after some regular transforms like log or sqrt. After binning, we do see differences of Grosses in different bins. See plots below.
```{r}
#summary dfn2$imdbRating
table(dfn2o$imdbRating)
#summary dfn2$Runtime
table(dfn2o$Runtime)
#plot Gross~imdbRating after binning
ggplot(dfn2o,aes(imdbRating,Gross,fill=imdbRating))+
  geom_boxplot()+coord_flip()+
  ggtitle("Gross vs imdbRating")
#plot Gross~Runtime after binning
ggplot(dfn2o,aes(Runtime,Gross,fill=Runtime))+
  geom_boxplot()+coord_flip()+
  ggtitle("Gross vs Runtime")
```

In task 1, the average train errors are around 96 M, and the average test errors are around 89 M.
After transformations in task 2, the average train errors dropped to around 91 M, and the average test errors dropped to around 82 M.

```{r}
#print average train rmse
average_rmse2$rmse[average_rmse2$type=="train"]
#print average test rmse 
average_rmse2$rmse[average_rmse2$type=="test"]
```


## 3. Non-numeric variables

Write code that converts genre, actors, directors, and other categorical variables to columns that can be used for regression (e.g. binary columns as you did in Project 1). Also process variables such as awards into more useful columns (again, like you did in Project 1). Now use these converted columns only to build your next model.

```{r}
# TODO: Build & evaluate model 3 (converted non-numeric variables only)
#make genres to binary columns
df3=df
genres=trimws(unlist(strsplit(df3$Genre,",")))
genreDict=unique(genres)

for(genre in genreDict){
  genreCol=data.frame(as.numeric(grepl(genre,df3$Genre)))
  names(genreCol)=genre
  df3=cbind(df3,genreCol)
}
df3$Genre=NULL
#make top 10 actors to binary columns
actors=trimws(unlist(strsplit(df3$Actors,",")))
topactorDict=names(head(sort(table(actors),decreasing = T),10))
for(actor in topactorDict){
  actorCol=data.frame(as.numeric(grepl(actor,df3$Actors)))
  names(actorCol)=actor
  df3=cbind(df3,actorCol)
}
df3$Actors=NULL
#make top 10 directors to binary columns
directors=trimws(unlist(strsplit(df3$Director,",")))
topdirectorDict=names(head(sort(table(directors),decreasing = T),10))
for(director in topdirectorDict){
  directCol=data.frame(as.numeric(grepl(director,df3$Director)))
  names(directCol)=director
  df3=cbind(df3,directCol)
}
df3$Director=NULL

#make wins binary columns
convert_wins=function(award){
  num_win=as.numeric(regmatches(award,regexec("(\\d+) *win",award))[[1]][2])
  num_won=as.numeric(regmatches(award,regexec("[Ww]on *(\\d+)",award))[[1]][2])
  sum(num_win,num_won,na.rm = T) #if NA, will set NA to 0
}

df3$Win=sapply(df3$Awards,convert_wins)
plot(Gross~Win,df3)
#cut wins to five classes

df3$WinRank[df3$Win==0]='NoWin'
df3$WinRank[df3$Win==1]='OneWin'
df3$WinRank[df3$Win==2|df3$Win==3]='TwoOrThreeWins'
df3$WinRank[df3$Win>3 & df3$Win<=10 ]='Fourto10Wins'
df3$WinRank[df3$Win>10]='MoreThan10Wins'

winlist=c('NoWin','OneWin','TwoOrThreeWins','Fourto10Wins','MoreThan10Wins')
for(winrank in winlist){
  winrankCol=data.frame(as.numeric(grepl(winrank,df3$WinRank)))
  names(winrankCol)=winrank
  df3=cbind(df3,winrankCol)
}
df3$WinRank=NULL
df3$Win=NULL
#make nominations binary columns:
convert_nominations=function(award){
  num_nominated=as.numeric(regmatches(award,regexec("Nominated *for *(\\d+)",award))[[1]][2])
  num_nomination=as.numeric(regmatches(award,regexec("(\\d+) *nomi",award))[[1]][2])
  sum(num_nominated,num_nomination,na.rm = T)#if NA, will set NA to 0
}
df3$Nomination=sapply(df3$Awards,convert_nominations)
#cut wins to six classes

df3$NomiRank[df3$Nomination==0]='NoNomi'
df3$NomiRank[df3$Nomination==1]='OneNomi'
df3$NomiRank[df3$Nomination==2|df3$Nomination==3]='TwoOrThreeNomi'
df3$NomiRank[df3$Nomination>3 & df3$Nomination<=7 ]='Fourto7Nomi'
df3$NomiRank[df3$Nomination>7 & df3$Nomination<=20]='Eightto20Nomi'
df3$NomiRank[df3$Nomination>20]='MoreThan20Nomi'

nomilist=c('NoNomi','OneNomi','TwoOrThreeNomi','Fourto7Nomi','Eightto20Nomi','MoreThan20Nomi')
for(nomirank in nomilist){
  nomirankCol=data.frame(as.numeric(grepl(nomirank,df3$NomiRank)))
  names(nomirankCol)=nomirank
  df3=cbind(df3,nomirankCol)
}
df3$Nomination=NULL
df3$NomiRank=NULL

df3$Awards=NULL

##make Rated to binary columns
unique(df3$Rated)
df3$Rated[df3$Rated=="NOT RATED" | df3$Rated=="UNRATED" |df3$Rated=="N/A"] = "UNKNOWN"
ratedlist=unique(df3$Rated)
for(rated in ratedlist){
  ratedCol=data.frame(as.numeric(grepl(rated,df3$Rated)))
  names(ratedCol)=rated
  df3=cbind(df3,ratedCol)
}
df3$Rated=NULL


#Deal with languages

df3$NumLang=sapply(df$Language,function(language) if(is.na(language)) NA else length(strsplit(language,",")[[1]]))
table(df3$NumLanguage)
df3$NumLanguage[df3$NumLang==1]="OneLanguage"
df3$NumLanguage[df3$NumLang==2]="TwoLanguage"
df3$NumLanguage[df3$NumLang==3]="ThreeLanguage"
df3$NumLanguage[df3$NumLang>=4]="MoreThan3Language"

numlanglist=c("OneLanguage","TwoLanguage","ThreeLanguage","MoreThan3Language")

for(numlang in numlanglist){
  numlangCol=data.frame(numl=as.numeric(grepl(numlang,df3$NumLanguage)))
  names(numlangCol)=numlang
  df3=cbind(df3,numlangCol)  
}
df3$NumLanguage=NULL
df3$Language=NULL
df3$NumLang=NULL
#make top 20 countries to binary columns
countries=trimws(unlist(strsplit(df3$Country,",")))
topcountryDict=names(head(sort(table(countries),decreasing = T),20))
for(country in topcountryDict){
  countryCol=data.frame(as.numeric(grepl(country,df3$Country)))
  names(countryCol)=country
  df3=cbind(df3,countryCol)
}
df3$Country=NULL

#make top 20 productions to binary columns
productions=trimws(unlist(strsplit(df3$Production,",")))
topproductionDict=names(head(sort(table(productions),decreasing = T),20))
for(production in topproductionDict){
  productionCol=data.frame(as.numeric(grepl(production,df3$Production)))
  names(productionCol)=production
  df3=cbind(df3,productionCol)
}
df3$Production=NULL

#use only converted columns to make models to predict Gross
df3_m=df3[,26:ncol(df3)]

#let's if we need to remove any columns
summary(lm(Gross~.,df3_m))
#we have to remove redundant columns
#remove "Joel Coen","NC-17","MoreThan10Wins","MoreThan20Nomi","MoreThan3Language" columns
df3_m=df3_m[,-c(40,49,55,61,66)]
```

```{r,warning=FALSE}
N=dim(df3_m)[1]
get_rmse_list3=function(df,prop_to_train,N){#helper funciton in task 3
  rmse_list=c()
  prop_to_train_list=c()
  type=c()
  set.seed(200)
  seeds=sample(10000,10)
  for( seed in seeds){
    set.seed(seed)
    to_train=sample(N,prop_to_train*N)
    train=df[to_train,]
    test=df[-to_train,]
    #make models
    model=lm(Gross~.,data = train)
    #make predictions and compute RMSEs
    pred_trainGross=predict(model,train[,-1])
    rmse_train=sqrt(mean((pred_trainGross-train[,1])^2))
    
    pred_testGross=predict(model,test[,-1])
    rmse_test=sqrt(mean((pred_testGross-test[,1])^2))
    #record results
    rmse_list=c(rmse_list,rmse_train,rmse_test)
    prop_to_train_list=c(prop_to_train_list,prop_to_train,prop_to_train)
    type=c(type,"train","test")
  }
  data.frame(prop_to_train=prop_to_train_list,type=type,rmse=rmse_list)
}
rmse_df3=get_rmse_list3(df3_m,0.05,N) #start iterations to compute RMSEs
for (prop_to_train in (2:19)*0.05){
  rmse_df3=rbind(rmse_df3,get_rmse_list3(df3_m,prop_to_train,N))
}
average_rmse3=aggregate(rmse_df3["rmse"],by=rmse_df3[c("type","prop_to_train")],function(X) c(mean=mean(X),sd=sd(X),N=length(X))) #summary RMSEs results
average_rmse3=do.call(data.frame,average_rmse3)
names(average_rmse3)[3:5]=c("rmse","sd","N")
average_rmse3$se=average_rmse3$sd/sqrt(average_rmse3$N)
#plot RMSEs in TASK3
ggplot(average_rmse3,aes(prop_to_train,rmse,group=type,col=type))+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin=rmse-se,ymax=rmse+se),width=0.02)+
  ylab("RMSE")+
  theme(axis.text.x = element_text(angle=90, hjust=1))+
  ggtitle("Task 3. RMSE over in and out of sample by train size")
```


**Q**: Explain which categorical variables you used, and how you encoded them into features.

**A**: 
1).make genres to binary columns
how encoded genres to features: 
First gain the all unique genres list, iterates the list to see if each row of Genre column has the genre appeared(1 for present, and 0 for not present). So we get 23 new genres columns. Then, the orignial Genre column is removed.
See detail codes from lines 380-390 in this rmd file.

2) make top 10 actors to binary columns
how encoded top 10 actors to features: 
First gain all actors appeared in the Actors column in a list. Table the list and get the actor names with top 10 occurance in the summary table. Then iterate top 10 actors to see if each row of Actors column has the actor appeared(1 for present, and 0 for not present). So we get 10 new actors columns. Then, the original Actors column is removed.
See detail codes from lines 391-399.

3) make top 10 directors to binary columns
how encoded top 10 directors to features: 
First gain all directors appeared in the Director column in a list. Table the list and get the director names with top 10 occurance in the summary table. Then iterate top 10 directors to see if each row of Director column has the director appeared(1 for present, and 0 for not present). So we get 10 new directors columns. Then, the original Director column is removed.

See detail codes from lines 400-408.

4) extract number of wins and nominations from Awards column and cut them into 5 and 6 classes respectively.
how encoded Awards to features:
In Awards column, first use regex to find number of wins and nominations for each row, and then bin them into several classes to make two new columns WinRank and NomiRank. See detail summary of two columns below:
```{r}
table(df3$WinRank)
table(df3$NomiRank)
```

Then remove Awards column.

Detail codes can be seen in lines 410-442.

5) make Rated to binary columns
how encoded Rated to features:
First check how many types of Rated in the column. combine "NOT RATED", "UNRATED","N/A" to a same type "UNKNOWN". Now we have the list of Rated.
Iterate the list to see if each row has the type of Rated appeared (1 for present, and 0 for absent). Now we have 7 new Rated columns. Then set original Rated column to NULL.

See detailed codes from line 444 to 453.

6) Deal with languages
How code Language to features:
First iterate each row of column Language to count how many languages appeared and assign the counts to a new column "NumLanguage". Make rows with more than 3 languages to a same class "MoreThan3". Make the column as factor type. So we have four levels in the column. See table below. Then remove Language column. Detail code can be seen from line 456 to 462.
```{r}
table(df3$NumLanguage)
```

7) make top 20 countries to binary columns
How code top 20 countries to features:

First gain all countries appeared in the Country column in a list. Table the list and get the country names with top 20 occurances in the summary table. Then iterate top 20 countries to see if each row of Country column has the country appeared(1 for present, and 0 for not present). So we get 20 new countries columns. Then, the original Country column is removed.

Detail codes can be seen from line 464 to 472.

8) make top 20 productions to binary columns
How code top 20 productions to features:

First gain all productions appeared in the Production column in a list. Table the list and get the production names with top 20 occurances in the summary table. Then iterate top 20 productions to see if each row of Production column has the production appeared(1 for present, and 0 for not present). So we get 20 new productions columns. Then, the original Production column is removed.

Detail codes can be seen from line 474 to 482.


In summary,  train RMSEs are around 132M and test RMSEs are around 122M in task 3. 
```{r}
#print average train RMSEs
average_rmse3$rmse[average_rmse3$type=="train"]

#print average test RMSEs 
average_rmse3$rmse[average_rmse3$type=="test"]
```


## 4. Numeric and categorical variables

Try to improve the prediction quality as much as possible by using both numeric and non-numeric variables from **Tasks 2 & 3**.

```{r,warning=FALSE}
# TODO: Build & evaluate model 4 (numeric & converted non-numeric variables)
#combine numeric variables from task 2 and converted variables from task 3
df4=cbind(dfn2,df3_m[,-1])
N=dim(df4)[1]
#helper function to iterate  10 times for computing RMSEs given a train size in task 4
get_rmse_list4=function(df,prop_to_train,N){
  #lists to record results
  rmse_list=c()
  prop_to_train_list=c()
  type=c()
  #set random seeds for good comparisions between tasks
  set.seed(200)
  seeds=sample(10000,10)
  #10 iterations
  for(seed in seeds){
    set.seed(seed)
    #data partitions
    to_train=sample(N,prop_to_train*N)
    train=df[to_train,]
    test=df[-to_train,]
    #build linear regression model
    model=lm(Gross~.,data = train)
    #make predicitons and compute RMSE
    pred_trainGross=(predict(model,train[,-1]))^2
    rmse_train=sqrt(mean((pred_trainGross-(train[,1])^2)^2))
    
    pred_testGross=(predict(model,test[,-1]))^2
    rmse_test=sqrt(mean((pred_testGross-(test[,1])^2)^2))
    #record results
    rmse_list=c(rmse_list,rmse_train,rmse_test)
    prop_to_train_list=c(prop_to_train_list,prop_to_train,prop_to_train)
    type=c(type,"train","test")
  }
  data.frame(prop_to_train=prop_to_train_list,type=type,rmse=rmse_list)
}
#iterate train size to compute RMSEs
rmse_df4=get_rmse_list4(df4,0.05,N)
for (prop_to_train in (2:19)*0.05){
  rmse_df4=rbind(rmse_df4,get_rmse_list4(df4,prop_to_train,N))
}
#summary RMSEs results to get mean and se
average_rmse4=aggregate(rmse_df4["rmse"],by=rmse_df4[c("type","prop_to_train")],function(X) c(mean=mean(X),sd=sd(X),N=length(X)))
average_rmse4=do.call(data.frame,average_rmse4)
names(average_rmse4)[3:5]=c("rmse","sd","N")
average_rmse4$se=average_rmse4$sd/sqrt(average_rmse4$N)

#plot RMSE results in task 4
ggplot(average_rmse4,aes(prop_to_train,rmse,group=type,col=type))+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin=rmse-se,ymax=rmse+se),width=0.02)+
  ylab("RMSE")+
  scale_y_continuous()+
  theme(axis.text.x = element_text(angle=90, hjust=1))+
  ggtitle("Task 4. RMSE over in and out of sample by train size")
```

After Task 4, Train RMSEs drop to around 85 M, and test RMSEs drop to around 77M.
```{r}
#print average train RMSEs
average_rmse4$rmse[average_rmse4$type=="train"]

#print average test RMSEs 
average_rmse4$rmse[average_rmse4$type=="test"]
```


## 5. Additional features

Now try creating additional features such as interactions (e.g. `is_genre_comedy` x `is_budget_greater_than_3M`) or deeper analysis of complex variables (e.g. text analysis of full-text columns like `Plot`).

```{r}
# TODO: Build & evaluate model 5 (numeric, non-numeric and additional features)
#label two clusters of tomatoUserReviews for later interation term in linear regression
df5=df4
df5$tomatoUserReviews=sqrt(df4$tomatoUserReviews)

df5$tomatoreviewrank[df5$tomatoUserReviews<=3000]=0
df5$tomatoreviewrank[df5$tomatoUserReviews>3000]=1

```

```{r}
f <- as.formula(Gross~.^2)
x <- model.matrix(f, df5)
cvgnetfit=cv.glmnet(x,df5$Gross,family="gaussian",type.measure = "mse",alpha=1)
plot(cvgnetfit)
tmp_coeffs <- coef(cvgnetfit, s = "lambda.min")
termname=rownames(tmp_coeffs)
goodterms=termname[which(tmp_coeffs != 0)][-1]
fml=formula(paste0("Gross~",paste0(goodterms,collapse = "+")))
```


```{r,warning=FALSE}
#helper function to iterate  10 times for computing RMSEs given a train size in task 5
get_rmse_list5=function(df,prop_to_train,N){
  #lists to record results
  rmse_list=c()
  prop_to_train_list=c()
  type=c()
  #set random seeds for good comparisons between taskes
  set.seed(200)
  seeds=sample(10000,10)
  #10 iterations 
  for( seed in seeds){
    set.seed(seed)
    #data partition
    to_train=sample(N,prop_to_train*N)
    train=df[to_train,]
    test=df[-to_train,]
    #build linear regression models with interaction terms
    #model=lm(Gross~.+Budget:imdbRating+Budget:tomatoUserRating+Budget:imdbVotes+imdbVotes:imdbRating+tomatoUserReviews:tomatoreviewrank,train)
    model=lm(fml,train)
    #make predicitons and compute RMSEs
    pred_trainGross=(predict(model,train[,-1]))^2
    rmse_train=sqrt(mean((pred_trainGross-(train[,1])^2)^2))
    
    pred_testGross=(predict(model,test[,-1]))^2
    rmse_test=sqrt(mean((pred_testGross-(test[,1])^2)^2))
    #record results
    rmse_list=c(rmse_list,rmse_train,rmse_test)
    prop_to_train_list=c(prop_to_train_list,prop_to_train,prop_to_train)
    type=c(type,"train","test")
  }
  data.frame(prop_to_train=prop_to_train_list,type=type,rmse=rmse_list)
}
#start iterations on train size to compute RMSEs
rmse_df5=get_rmse_list5(df5,0.05,N)
for (prop_to_train in (2:19)*0.05){
  rmse_df5=rbind(rmse_df5,get_rmse_list5(df5,prop_to_train,N))
}
#summary RMSEs and get mean and se
average_rmse5=aggregate(rmse_df5["rmse"],by=rmse_df5[c("type","prop_to_train")],function(X) c(mean=mean(X),sd=sd(X),N=length(X)))
average_rmse5=do.call(data.frame,average_rmse5)
names(average_rmse5)[3:5]=c("rmse","sd","N")
average_rmse5$se=average_rmse5$sd/sqrt(average_rmse5$N)
#plot RMSEs in task 5
task5plot1=ggplot(average_rmse5,aes(prop_to_train,rmse,group=type,col=type))+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin=rmse-se,ymax=rmse+se),width=0.02)+
  ylab("RMSE")+
  theme(axis.text.x = element_text(angle=90, hjust=1))+
  ggtitle("Task 5. RMSE over in and out of sample by train size")
task5plot1
#zoom in plot of task5
task5plot1+ coord_cartesian(ylim= c(0,150000000))
```

**Q**: Explain what new features you designed and why you chose them.

**A**: 
I add five interaction terms in the models used in Task 5. Here are the interaction terms: Budget:imdbRating,Budget:tomatoUserRating, Budget:imdbVotes, imdbVotes:imdbRating, tomatoUserReviews:tomatoreviewrank.

First reason, I add them is that when I stepwise add those interaction terms, I see drops in train RMSE by models from full size of data. And also I see the obvious significance effect of the interaction term in the model. Check the full model summary below.

```{r}
s5=summary(lm(fml,df5))
s5
```

And also I see increased Adjusted R-squared from model 4 to model 5 based on full data. See below.
```{r}
s4=summary(lm(Gross~.,df4))
#s4$adj.r.squared
print(s4$adj.r.squared)
#s5$adj.r.squared
print(s5$adj.r.squared)
```


Second, it makes sense to include them in the model. Such as tomatoUserReviews:tomatoreviewrank.  See plot below. We know there are two clusters of points, and for each cluster of data, the slope and intercept should be different. After including tomatoreviewrank and tomatoUserReviews:tomatoreviewrank, the intercept and slope for two clusters are adjusted accordingly. 

```{r}
#based on plot below, I included interation term of tomatoUserReviews:tomatoreviewrank
ggplot(data=df5,aes(tomatoUserReviews,Gross,color=as.factor(tomatoreviewrank)))+
  geom_point()+ggtitle("Gross vs tomatoUserReview by tomatoreviewrank")
```


After Task 5, train RMSEs drop to around 56 M and test RMSEs drop to around 66 M.
```{r}
#print average train RMSEs
average_rmse5$rmse[average_rmse5$type=="train"]

#print average test RMSEs 
average_rmse5$rmse[average_rmse5$type=="test"]
```




#summary RMSE evaluation results from task 1 to task 5
We see the trend of drop of RMSEs from task 1 to task 2, from task 2/task 3 to task 4, and from task 4 to task 5.
```{r}
average_rmse$task="1"
average_rmse2$task="2"
average_rmse3$task="3"
average_rmse4$task="4"
average_rmse5$task="5"
average_rmse_all=rbind(average_rmse,average_rmse2,average_rmse3,average_rmse4,average_rmse5)

summaryplot1=ggplot(average_rmse_all,aes(prop_to_train,rmse,group=task,col=task))+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin=rmse-se,ymax=rmse+se),width=0.02)+
  ylab("RMSE")+
  theme(axis.text.x = element_text(angle=90, hjust=1))+
  facet_grid(.~type)+
  ggtitle("Summary Task 1 to 5. RMSE over in and out of sample by train size")
summaryplot1
#let's zoom in the summary plot
summaryplot1+coord_cartesian(ylim= c(0,300000000))
```

Here is the summary of  the adjusted R squared from task 1 to task 5.
We see improvement from task 1 to task 2, from task 2/task 3 to task 4, from task 4 to task 5.
```{r}
s1=summary(lm(Gross~.,dfn))
s2=summary(lm(Gross~.,dfn2))
s3=summary(lm(Gross~.,df3_m))
R2adj=c(s1$adj.r.squared,s2$adj.r.squared,s3$adj.r.squared,s4$adj.r.squared,s5$adj.r.squared)
Tasks=c("Task 1","Task 2","Task 3","Task 4", "Task 5")
radj_df=data.frame(R2adj=R2adj,Task=Tasks)
ggplot(radj_df,aes(Task,R2adj,fill=Task))+geom_bar(stat="identity")+
  ylab("adjusted R squared")+
  geom_text(aes(label = round(R2adj,digits = 4),y=R2adj+0.05), size = 4)+
  ggtitle("adjusted R^2 by Tasks")
```


